---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

![stability-unstable](https://img.shields.io/badge/stability-unstable-yellow.svg)
[![Travis-CI Build Status](https://travis-ci.org/poissonconsulting/jmbr.svg?branch=master)](https://travis-ci.org/poissonconsulting/jmbr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/jmbr?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/jmbr)
[![codecov](https://codecov.io/gh/poissonconsulting/jmbr/branch/master/graph/badge.svg)](https://codecov.io/gh/poissonconsulting/jmbr)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)

# jmbr

## Introduction

`jmbr` (pronounced jimber) is an R package to facilitate analyses using Just Another Gibbs Sampler (JAGS).
It is part of the [mbr](https://github.com/poissonconsulting/mbr) family of packages.

## Demonstration

```{r, message = FALSE}
library(magrittr)
library(ggplot2)
library(jmbr)
```

```{r}
# define model in JAGS language
model <- model("model {
  alpha ~ dnorm(0, 10^-2)
  beta1 ~ dnorm(0, 10^-2)
  beta2 ~ dnorm(0, 10^-2)
  beta3 ~ dnorm(0, 10^-2)

  log_sDispersion ~ dnorm(0, 10^-2)

  log(sDispersion) <- log_sDispersion

  for (i in 1:length(Pairs)) {
    log(ePairs[i]) <- alpha + beta1 * Year[i] + beta2 * Year[i]^2 + beta3 * Year[i]^3
    eDispersion[i] ~ dgamma(1 / sDispersion^2, 1 / sDispersion^2)
    Pairs[i] ~ dpois(ePairs[i] * eDispersion[i])
  }
}")

# add R code to calculate derived parameters
model %<>% update_model(new_expr = "
for (i in 1:length(Pairs)) {
  prediction[i] <- exp(alpha + beta1 * Year[i] + beta2 * Year[i]^2 + beta3 * Year[i]^3)
}")

# define data types and center year
model %<>% update_model(select_data = list("Pairs" = integer(), "Year*" = integer()))

# analyse
analysis <- analyse(model, data = bauw::peregrine)
analysis %<>% reanalyse(rhat = 1.05)

coef(analysis)

plot(analysis)
```

```{r, message = FALSE}
# make predictions by varying year with other predictors held constant
year <- predict(analysis, new_data = "Year")

# plot those predictions
ggplot(data = year, aes(x = Year, y = estimate)) +
  geom_point(data = bauw::peregrine, aes(y = Pairs)) +
  geom_line() +
  geom_line(aes(y = lower), linetype = "dotted") +
  geom_line(aes(y = upper), linetype = "dotted") +
  expand_limits(y = 0)
```

## Installation

To install from GitHub
```
# install.packages("devtools")
devtools::install_github("poissonconsulting/jmbr")
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/jmbr/issues).

[Pull requests](https://github.com/poissonconsulting/jmbr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Inspiration

- [jaggernaut](https://github.com/poissonconsulting/jaggernaut)
